---
title: "HW3"
author: "Chris Hill"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
load("hw2_list.rda")
str(bp_list)
```

# Task 1: Control Flow Practice

### (1) Create the status Column

```{r}
bp_list <- lapply(bp_list, function(df) {
  df$status <- character(nrow(df))
  return(df)
})

str(bp_list)
```

### (2) Update status for the treatment Data Frame

```{r}
for (i in seq_len(nrow(bp_list$treatment))) {
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  } else if (bp_list$treatment$post_bp[i] <= 130) {
    bp_list$treatment$status[i] <- "borderline"
  } else {
    bp_list$treatment$status[i] <- "high"
  }
}

bp_list$treatment |> head() |> knitr::kable()
```

### (3) Update status for the placebo Data Frame

```{r}
for (i in seq_len(nrow(bp_list$placebo))) {
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  } else if (bp_list$placebo$post_bp[i] <= 130) {
    bp_list$placebo$status[i] <- "borderline"
  } else {
    bp_list$placebo$status[i] <- "high"
  }
}

bp_list$placebo |> head() |> knitr::kable()
```

# Task 2: Function Writing

### (1) Define the Function

```{r}
calculate_stats <- function(df_list, stat = "mean") {
  # Retrieve the function corresponding to the statistic
  my_fun <- get(stat)
  
  # Calculate the statistics for each column in both data frames
  treatment_pre <- my_fun(df_list$treatment$pre_bp, na.rm = TRUE)
  treatment_post <- my_fun(df_list$treatment$post_bp, na.rm = TRUE)
  treatment_diff <- my_fun(df_list$treatment$diff_bp, na.rm = TRUE)
  
  placebo_pre <- my_fun(df_list$placebo$pre_bp, na.rm = TRUE)
  placebo_post <- my_fun(df_list$placebo$post_bp, na.rm = TRUE)
  placebo_diff <- my_fun(df_list$placebo$diff_bp, na.rm = TRUE)
  
  # Create a named vector to store the results
  stats_vector <- c(
    treatment_pre, treatment_post, treatment_diff,
    placebo_pre, placebo_post, placebo_diff
  )
  
  # Create meaningful names for the vector
  names(stats_vector) <- c(
    paste0(stat, "_treatment_pre"),
    paste0(stat, "_treatment_post"),
    paste0(stat, "_treatment_diff"),
    paste0(stat, "_placebo_pre"),
    paste0(stat, "_placebo_post"),
    paste0(stat, "_placebo_diff")
  )
  
  # Return the named vector
  return(stats_vector)
}
```

### (2) Apply the Function to the bp_list

*Apply the function to list of data frames from previous task. Use it without specifying the statistic, then with specifying your statistic as "var", "sd", "min", and "max"*

#### Mean (without specifying statistic)

```{r}
result_mean <- calculate_stats(bp_list)
knitr::kable(result_mean)
```

#### Variance

```{r}
result_var <- calculate_stats(bp_list, stat = "var")
knitr::kable(result_var)
```

#### Standard Deviation

```{r}
result_sd <- calculate_stats(bp_list, stat = "sd")
knitr::kable(result_sd)
```

#### Min

```{r}
result_min <- calculate_stats(bp_list, stat = "min")
knitr::kable(result_min)
```

#### Max

```{r}
result_max <- calculate_stats(bp_list, stat = "max")
knitr::kable(result_max)
```

